Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-03-17 paper/paper-MinD.tex
switched changes over to plos.tex, adjusted figures to fit better Tested everything on crystal.

\fixme{
  First referee:
  \begin{enumerate}
  \item \blue{David} Add discussion of ``platform'' --- wrote our own code,
    publically available (cite github?).
  \item \blue{David} Make 3D geometries available? Cite that the publically
    available code can generate them?
  \item Claims differential equations aren't needed.
  \item Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item Lacks diffusion on the membrane. Supported by Meacci 2006,
      which experimentally tries to measure diffusion on membrane, has
      a tough time, but shows that definitely diffuses.
    \item Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item (Still debated) No evidence that DE recruits to membrane.
    \item Works in only a limited part of parameters phase space, and
      could not account for important things like
      temperature-independence.
    \end{enumerate}
  \item \blue{David} - Height should be bigger to compare
    with experiment.  0.4-0.6 microns.  \textcolor{blue}{HIGH
      PRIORITY!!!  to figure this out}

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$,
respectively. \blue{(1,2)} We have chosen to write our own simulation
platform and membrane creation tools from scratch, instead of using
existing available software used in previous studies\fixme{cite
  previous studies}. We believe there is value in re-creating results
with alternative software.  Our simulations take place within a
three-dimensional Cartesian grid that has a grid spacing of
.05\micron, and our cell shapes are specified as the zero of an
analytic function which can generate the geometry presented in this
paper.  Our code is publically available on github.com~\fixme{cite
  github account in references?}

@@

It is clear from Fig.~\ref{image-p} that the stochastic simulation
results in a protein distribution that is less spatially regular than
the deterministic model predicts.  In fact, when viewed as a movie
(see supplementary material), the stochastic model exhibits a ``starry
night'' effect, as recruitment leads to clusters of MinD forming on
the wall and then subsequently dissipating.  \fixme{Talk about
  membrane diffusion blurring out the starry night.}  This naturally
leads to irregularity in the temporal periodicity of the system, which
we quantify by examining the temporal correlation function of the
total MinD found in two opposite polar regions.
%
This correlation function, which is displayed in Fig.~\ref{corr-pill},
is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

@@

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation, with the
exception of shape \emph{B}, which develops weak intermediate maxima
on the right-hand edge of the cell.  From these results, we conclude
that the deterministic model is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

>>>>

\fixme{
  First referee:
  \begin{enumerate}
  \item \blue{David} Add discussion of ``platform'' --- wrote our own code,
    publically available (cite github?).
  \item \blue{David} Make 3D geometries available? Cite that the publically
    available code can generate them?
  \item Claims differential equations aren't needed.
  \item Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item \blue{David.  I've added a paragraph discussing the starry
      night in the context of the membrane diffusion.  I didn't say
      anything about a one half period allowing diffusion to smear an
      entire pole, because it seems to contradict the next paragraph.
      What do you think?} Lacks diffusion on the membrane. Supported
      by Meacci 2006, which experimentally tries to measure diffusion
      on membrane, has a tough time, but shows that definitely
      diffuses.
    \item Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item (Still debated) No evidence that DE recruits to membrane.
    \item Works in only a limited part of parameters phase space, and
      could not account for important things like
      temperature-independence.
    \end{enumerate}
  \item \blue{David} - Height should be bigger to compare
    with experiment.  0.4-0.6 microns.  \textcolor{blue}{HIGH
      PRIORITY!!!  to figure this out}

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$,
respectively.  We have chosen to write our own simulation
platform and membrane creation tools from scratch, instead of using
existing available software used in previous studies\fixme{cite
  previous studies} (we believe there is value in re-creating results
with alternative software).  Our simulations take place within a
three-dimensional Cartesian grid that has a grid spacing of
.05\micron, and our cell shapes are specified as the zero of an
analytic function which can generate the geometry presented in this
paper.  Our code is publically available on github.com~\fixme{cite
  github account in references?}

@@

\blue{(5a)} Both Fig.~\ref{image-p} and its associated movies (see
supplementary material), show that the stochastic model exhibits a
``starry night'' effect characterized by spatially fixed points of
protein density build up, as recruitment leads to clusters of MinD
forming on the wall and then subsequently dissipating.  We attribute
this effect to the fact that the Huang model does not include
diffusion along the membrane, which has been experimentally verified
yet found to be two orders of magnitude less than diffusion in the
cytoplasm~\cite{meacci2005min}. Even ignoring this effect, however, it
is clear from Fig.~\ref{image-p} that the stochastic simulation
results in a protein distribution that is less spatially regular than
the deterministic model predicts.  This naturally leads to
irregularity in the temporal periodicity of the system, which we
quantify by examining the temporal correlation function of the total
MinD found in two opposite polar regions.
%
This correlation function, which is displayed in Fig.~\ref{corr-pill},
is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

@@

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation.  From these
results, we conclude that the deterministic model is inadequate to
explain experimental observations of the locations of density maxima
of the MinD protein.
